---
title: "CUT&RUN Peak Calling: MACS2 and SEACR"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: true
    toc_float: true
    #number_sections: false
    df_print: paged
    code_folding: show
editor_options: 
  chunk_output_type: console
---

<!-- <style> -->
<!-- body{text-align: justify} -->
<!-- pre code, pre, code { -->
<!--   white-space: pre !important; -->
<!--   overflow-x: scroll !important; -->
<!--   word-break: keep-all !important; -->
<!--   word-wrap: initial !important; -->
<!-- } -->
<!-- </style> -->


```{r setup, include=FALSE, dev="CairoPNG"}
knitr::opts_chunk$set(echo = TRUE, dev="CairoPNG", out.width = "100%")#, cache=TRUE)
```

```{r echo=FALSE, eval=TRUE, message=FALSE, warning=FALSE}
library(GenomicRanges)
library(dplyr)
library(data.table)
library(ggplot2)
library(viridis)
library(readxl)
library(pheatmap)
library(RColorBrewer)
library(tidyr)
library(gridExtra)
library(corrplot)
library(ggrepel)
library(tibble)
```


## I. Keeping duplicates

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
inPath = "CART_CUTRUN_Project/results/CUTRUN/process/"
outPath = "CART_CUTRUN_Project/results/CUTRUN/analysis/"
fileList = fread("CART_CUTRUN_Project/meta/Histone_IgG_pairing.csv")$histone

convertGR = function(peak){
    peak.gr = GRanges(seqname = peak$V1, IRanges(peak$V2, peak$V3), strand = "*")
    return(peak.gr)
}
peakN = c()
peakLen = c()
peakCov = c()
peakOverlap = c()
for(file in fileList){
    print(file)
    macsPeak = fread(paste0(inPath, file, "/peakCalling/relax/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM")
    seacrControl = fread(paste0(inPath, file, "/peakCalling/relax/SEACR/seacr_control.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM")
    seacrTop = fread(paste0(inPath, file, "/peakCalling/relax/SEACR/seacr_top0.01.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM")


    #peakN = data.frame(peakN = c(macsPeak %>% nrow, seacrControl %>% nrow, seacrTop %>% nrow), type = c("MACS2", "SEACR_Control", "SEACR_Top1%"), info = file) %>% rbind(peakN, .)

    #peakLen = data.frame(peakLen = c(macsPeak$V3 - macsPeak$V2, seacrControl$V3 - seacrControl$V2, seacrTop$V3 - seacrTop$V2), type = c(rep("MACS2", nrow(macsPeak)), rep("SEACR_Control", nrow(seacrControl)), rep("SEACR_Top1%", nrow(seacrTop))), info = file) %>% rbind(peakLen, .)

    #peakCov = data.frame(peakLen = c(sum(macsPeak$V3 - macsPeak$V2), sum(seacrControl$V3 - seacrControl$V2), sum(seacrTop$V3 - seacrTop$V2)), type = c("MACS2", "SEACR_Control", "SEACR_Top1%"), info = file) %>% rbind(peakCov, .)

    macsPeak.gr = convertGR(macsPeak %>% top_n(nrow(seacrControl), wt = V9))
    seacrControl.gr = convertGR(seacrControl)
    seacrTop.gr = convertGR(seacrTop)

    macsSeacr = findOverlaps(macsPeak.gr, seacrControl.gr)
    seacrCT = findOverlaps(seacrControl.gr, seacrTop.gr)
    peakOverlap = data.frame(overlapN = c(macsSeacr@from %>% unique %>% length, macsSeacr@to %>% unique %>% length, seacrCT@from %>% unique %>% length, seacrCT@to %>% unique %>% length),
                             overlapProp =  c(macsSeacr@from %>% unique %>% length/length(macsPeak.gr) * 100, macsSeacr@to %>% unique %>% length/length(seacrControl.gr) * 100, seacrCT@from %>% unique %>% length/length(seacrControl.gr) * 100, seacrCT@to %>% unique %>% length/length(seacrTop.gr) *100), propLabel = c("MACS2vsSEACR/MACS2", "MACS2vsSEACR/SEACR", "SEACR_ControlvsTop/Control", "SEACR_ControlvsTop/Top"), numLabel = c("MACS2vsSEACR:MACS2", "MACS2vsSEACR:SEACR", "SEACR_ControlvsTop:Control", "SEACR_ControlvsTop:Top"), info = file) %>% rbind(peakOverlap, .)


    # peakOverlap = data.frame(overlapN = c(macsSeacr@from %>% unique %>% length, macsSeacr@to %>% unique %>% length, seacrCT@from %>% unique %>% length, seacrCT@to %>% unique %>% length), overlapProp =  c(macsSeacr@from %>% unique %>% length/nrow(macsPeak) * 100, macsSeacr@to %>% unique %>% length/nrow(seacrControl) * 100, seacrCT@from %>% unique %>% length/nrow(seacrControl) * 100, seacrCT@to %>% unique %>% length/nrow(seacrTop) *100), propLabel = c("MACS2vsSEACR/MACS2", "MACS2vsSEACR/SEACR", "SEACR_ControlvsTop/Control", "SEACR_ControlvsTop/Top"), numLabel = c("MACS2vsSEACR:MACS2", "MACS2vsSEACR:SEACR", "SEACR_ControlvsTop:Control", "SEACR_ControlvsTop:Top"), info = file) %>% rbind(peakOverlap, .)
}
save(peakN, peakCov, peakLen, peakOverlap, file = paste0(outPath, "RData/peakOverlap_withDuplicates.RData"))

load(file = paste0(outPath, "RData/peakOverlap_withDuplicates.RData"))
```
 
 - chrY and chrM are removed in this analysis.
 
 
### 1. Number of peaks

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=13, fig.height=9}
peakNData = separate(peakN, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

data = peakNData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP")
data %>% ggplot(aes(x = CellType, y = as.numeric(peakN), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
   geom_label_repel(data = data %>% filter(Experiment == "Input", HD %in% paste0("HD", c(1, 2, 3))), aes(label = paste0(Experiment, "_", HD)), position = position_jitterdodge()) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 15) +
    ylab("# of Peaks") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```

Zoom in y-scale:0 - 40,000
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=13, fig.height=9}
data = peakNData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP")
data %>% ggplot(aes(x = CellType, y = as.numeric(peakN), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    geom_label_repel(data = data %>% filter(Experiment == "Input", HD %in% paste0("HD", c(1, 2, 3))), aes(label = paste0(Experiment, "_", HD)), position = position_jitterdodge()) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 15) +
    ylab("# of Peaks") +
    xlab("") +
    ggtitle("Keeping Duplicates") +
  coord_cartesian(ylim = c(0, 40000))
```

- For MACS2, if set ```--call-summit``` it may give multiple summits for one large peak regions, namely call multiple subpeaks within one large peak if it has multiple summits.

- In this figure, ```--call-summit``` is not set.
 
### 2. peak length distribution

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
# peakLenData = separate(peakLen, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))
# 
# save(peakLenData, file = paste0(outPath, "RData/peakOverlap_peakLenData_withDuplicates.RData"))
load(file = paste0(outPath, "RData/peakOverlap_peakLenData_withDuplicates.RData"))

peakLenData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakLen), fill = CellType)) +
    geom_violin() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("Length of Peaks (bp)") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```

Zoom in y-scale: 0-15,000
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakLenData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakLen), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("Length of Peaks (bp)") +
    xlab("") +
    coord_cartesian(ylim = c(0, 15000)) +
    ggtitle("Keeping Duplicates")

```

 - H3K27me3 is expected to have wider peaks than H3K4me2.

### 3. peak coverage

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakCovData = separate(peakCov, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

peakCovData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakLen), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("Coverage of Peaks (bp)") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```

### 4. overlap between MACS and SEACR

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
peakOverlapData = separate(peakOverlap, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

peakOverlapData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(overlapProp), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~propLabel, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("% of Peaks Overlapped") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```

There are four categories (columns):

- (MACS2 overlapping SEACR)/top N peaks called by MACS2 * 100, where N = # of peaks called by SEACR using IgG as control.
- (MACS2 overlapping SEACR)/# of peaks called by SEACR using IgG as control * 100
- (SEACR using IgG as control overlapping SEACR not using IgG but take the top 1%)/SEACR using control * 100
- (SEACR using IgG as control overlapping SEACR not using IgG but take the top 1%)/SEACR top 1% * 100
 


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
peakOverlapData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(overlapN), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~numLabel, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("# of Peaks Overlapped") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```



### 5. Reproducibility across human donors

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
inPath = "CART_CUTRUN_Project/results/CUTRUN/process/"
outPath = "CART_CUTRUN_Project/results/CUTRUN/analysis/"

blackList = c("H3K27me3_CD8_CM_Input_HD1_rep1", "H3K27me3_CD8_CM_Input_HD3_rep2", "H3K27me3_CD8_N_Input_HD2_rep1", "H3K27me3_CD8_N_Input_HD3_rep2")
# fileList = fread("/fh/tempspace/gottardo_r/yezheng_working/meta/Histone_IgG_pairing.csv") %>% select(histone) %>% mutate(path = histone) %>% filter(! histone %in% blackList) %>% separate("histone", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") 
# fileListUni = fileList %>% select(Histone, CD, CellType, Experiment) %>% unique
# convertGR = function(peak){
#     peak.gr = GRanges(seqname = peak$V1, IRanges(peak$V2, peak$V3), strand = "*")
#     return(peak.gr)
# }
# 
# reprodD = c()
# for(i in 1:nrow(fileListUni)){
#   print(i)
#   hist = fileListUni[i, 1]
#   cd = fileListUni[i, 2]
#   cell = fileListUni[i, 3]
#   exp = fileListUni[i, 4]
#   hdList = fileList %>% filter(Histone == hist, CD == cd, CellType == cell, Experiment == exp)
#   for(p1 in 1:nrow(hdList)){
#       for(p2 in 1:nrow(hdList)){
#         if(p1 != p2){
#           
#           ## MACS2
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/relax/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/relax/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "MACS2") %>% rbind(reprodD)
#           
#            ## SEACR Control
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/relax/SEACR/seacr_control.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/relax/SEACR/seacr_control.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "SEACR_Control") %>% rbind(reprodD)
#     
#           p1N = length(peak1.gr)
#           p2N = length(peak2.gr)
#           peak1Con.gr = peak1.gr
#           peak2Con.gr = peak2.gr
#           
#           ## SEACR Top
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/relax/SEACR/seacr_top0.01.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/relax/SEACR/seacr_top0.01.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "SEACR_Top") %>% rbind(reprodD)
#           
#           ## SEACR Control in Top
#           peak1ConTop.gr = peak1Con.gr[findOverlaps(peak1Con.gr, peak1.gr)@from %>% unique]
#           peak2ConTop.gr = peak2Con.gr[findOverlaps(peak2Con.gr, peak2.gr)@from %>% unique]
#           overlap = findOverlaps(peak1ConTop.gr, peak2ConTop.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1ConTop.gr) * 100, peakN = length(peak1ConTop.gr), caller = "SEACR_ControlInTop") %>% rbind(reprodD)
#           
#           ## MACS2 - topN
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/relax/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% top_n(p1N, wt = V9) %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/relax/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% top_n(p2N, wt = V9) %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "MACS2(TopN=SEACR_Control)") %>% rbind(reprodD)
#           
#           
#           
#      
#           
#         }
#         
#       }
#     
#   }
#   
# }
# reprodD$caller = factor(reprodD$caller, levels = c("MACS2", "SEACR_Control", "SEACR_Top", "MACS2(TopN=SEACR_Control)", "SEACR_ControlInTop"))
# save(reprodD, file = paste0(outPath, "RData/reprodD_withDuplicates.RData"))


load(file = paste0(outPath, "RData/reprodD_withDuplicates.RData"))
```

#### 5.1 Variation of number of peaks among donors

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
reprodData = separate(reprodD, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

reprodData %>% select(-Replicate, -overlapHD, -overlapN, -overlapP) %>% unique %>% group_by(Histone, CD, CellType, Experiment, caller) %>% summarize(sd = sd(peakN)) %>%
     filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = sd, fill = CellType)) +
    geom_boxplot() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~caller, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("Standard Deviation of Number of Peaks being Called") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```

Zoom in y-scale: 0 - 10,000
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
reprodData %>% dplyr::select(-Replicate, -overlapHD, -overlapN, -overlapP) %>% unique %>% group_by(Histone, CD, CellType, Experiment, caller) %>% summarize(sd = sd(peakN)) %>%
     filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = sd, fill = CellType)) +
    geom_boxplot() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~caller, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("Standard Deviation of Number of Peaks being Called") +
    xlab("") +
    ggtitle("Keeping Duplicates") + 
    coord_cartesian(ylim = c(0,10000))
```

#### 5.2 Peak overlapping among donors

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
reprodData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(overlapP), fill = CellType)) +
    geom_boxplot() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~caller, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("Reproducibility (%)") +
    xlab("") +
    ggtitle("Keeping Duplicates")
```



## II. Removing duplicates

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
inPath = "CART_CUTRUN_Project/results/CUTRUN/process/"
outPath = "CART_CUTRUN_Project/results/CUTRUN/analysis/"
fileList = fread("CART_CUTRUN_Project/meta/Histone_IgG_pairing.csv")$histone

# convertGR = function(peak){
#     peak.gr = GRanges(seqname = peak$V1, IRanges(peak$V2, peak$V3), strand = "*")
#     return(peak.gr)
# }
# peakN = c()
# peakLen = c()
# peakCov = c()
# peakOverlap = c()
# for(file in fileList){
#     print(file)
#     macsPeak = fread(paste0(inPath, file, "/peakCalling/stringent/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM")
#     seacrControl = fread(paste0(inPath, file, "/peakCalling/stringent/SEACR/seacr_control.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM")
#     seacrTop = fread(paste0(inPath, file, "/peakCalling/stringent/SEACR/seacr_top0.01.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM")
#     
#     
#     peakN = data.frame(peakN = c(macsPeak %>% nrow, seacrControl %>% nrow, seacrTop %>% nrow), type = c("MACS2", "SEACR_Control", "SEACR_Top1%"), info = file) %>% rbind(peakN, .)
#     
#     peakLen = data.frame(peakLen = c(macsPeak$V3 - macsPeak$V2, seacrControl$V3 - seacrControl$V2, seacrTop$V3 - seacrTop$V2), type = c(rep("MACS2", nrow(macsPeak)), rep("SEACR_Control", nrow(seacrControl)), rep("SEACR_Top1%", nrow(seacrTop))), info = file) %>% rbind(peakLen, .)
#     
#     peakCov = data.frame(peakLen = c(sum(macsPeak$V3 - macsPeak$V2), sum(seacrControl$V3 - seacrControl$V2), sum(seacrTop$V3 - seacrTop$V2)), type = c("MACS2", "SEACR_Control", "SEACR_Top1%"), info = file) %>% rbind(peakCov, .)
#     
#     macsPeak.gr = convertGR(macsPeak %>% top_n(nrow(seacrControl), wt = V9))
#     seacrControl.gr = convertGR(seacrControl)
#     seacrTop.gr = convertGR(seacrTop)
#     
#     macsSeacr = findOverlaps(macsPeak.gr, seacrControl.gr)
#     seacrCT = findOverlaps(seacrControl.gr, seacrTop.gr)
#     peakOverlap = data.frame(overlapN = c(macsSeacr@from %>% unique %>% length, macsSeacr@to %>% unique %>% length, seacrCT@from %>% unique %>% length, seacrCT@to %>% unique %>% length), 
#                              overlapProp =  c(macsSeacr@from %>% unique %>% length/length(macsPeak.gr) * 100, macsSeacr@to %>% unique %>% length/length(seacrControl.gr) * 100, seacrCT@from %>% unique %>% length/length(seacrControl.gr) * 100, seacrCT@to %>% unique %>% length/length(seacrTop.gr) *100), propLabel = c("MACS2vsSEACR/MACS2", "MACS2vsSEACR/SEACR", "SEACR_ControlvsTop/Control", "SEACR_ControlvsTop/Top"), numLabel = c("MACS2vsSEACR:MACS2", "MACS2vsSEACR:SEACR", "SEACR_ControlvsTop:Control", "SEACR_ControlvsTop:Top"), info = file) %>% rbind(peakOverlap, .)
#     
#     #peakOverlap = data.frame(overlapN = c(macsSeacr@from %>% unique %>% length, macsSeacr@to %>% unique %>% length, seacrCT@from %>% unique %>% length, seacrCT@to %>% unique %>% length), overlapProp =  c(macsSeacr@from %>% unique %>% length/nrow(macsPeak) * 100, macsSeacr@to %>% unique %>% length/nrow(seacrControl) * 100, seacrCT@from %>% unique %>% length/nrow(seacrControl) * 100, seacrCT@to %>% unique %>% length/nrow(seacrTop) *100), propLabel = c("MACS2vsSEACR/MACS2", "MACS2vsSEACR/SEACR", "SEACR_ControlvsTop/Control", "SEACR_ControlvsTop/Top"), numLabel = c("MACS2vsSEACR:MACS2", "MACS2vsSEACR:SEACR", "SEACR_ControlvsTop:Control", "SEACR_ControlvsTop:Top"), info = file) %>% rbind(peakOverlap, .)
#    
# }
# save(peakN, peakCov, peakLen, peakOverlap, file = paste0(outPath, "RData/peakOverlap_rmDuplicates.RData"))

load(file = paste0(outPath, "RData/peakOverlap_rmDuplicates.RData"))
```

### 1. Number of peaks

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakNData = separate(peakN, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

peakNData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakN), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("# of Peaks") +
    xlab("") +
    ggtitle("Removing Duplicates")

    
```

Zoom in y-scale: 0 - 40,000

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakNData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakN), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("# of Peaks") +
    xlab("") +
    ggtitle("Removing Duplicates") +
  coord_cartesian(ylim = c(0, 40000))
```

### 2. peak length distribution

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakLenData = separate(peakLen, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

save(peakLenData, file = paste0(outPath, "RData/peakOverlap_peakLenData_rmDuplicates.RData"))
load(file = paste0(outPath, "RData/peakOverlap_peakLenData_rmDuplicates.RData"))

peakLenData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakLen), fill = CellType)) +
    geom_violin() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("Length of Peaks (bp)") +
    xlab("") +
    ggtitle("Removing Duplicates")
```

Zoom in y-scale:0-15,000
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakLenData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakLen), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("Length of Peaks (bp)") +
    xlab("") +
    coord_cartesian(ylim = c(0, 15000)) +
    ggtitle("Removing Duplicates")
```

### 3. peak coverage

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
peakCovData = separate(peakCov, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

peakCovData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(peakLen), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~type, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw() +
    ylab("Coverage of Peaks (bp)") +
    xlab("") +
    ggtitle("Removing Duplicates")
```

### 4. Overlap between MACS and SEACR

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
peakOverlapData = separate(peakOverlap, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

peakOverlapData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(overlapProp), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~propLabel, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("% of Peaks Overlapped") +
    xlab("") +
    ggtitle("Removing Duplicates")
```

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
peakOverlapData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(overlapN), fill = CellType)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~numLabel, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("# of Peaks Overlapped") +
    xlab("") +
    ggtitle("Removing Duplicates")
```

### 5. Reproducibility across human donors

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
inPath = "CART_CUTRUN_Project/results/CUTRUN/process/"
outPath = "CART_CUTRUN_Project/results/CUTRUN/analysis/"

# blackList = c("H3K27me3_CD8_CM_Input_HD1_rep1", "H3K27me3_CD8_CM_Input_HD3_rep2", "H3K27me3_CD8_N_Input_HD2_rep1", "H3K27me3_CD8_N_Input_HD3_rep2")
# fileList = fread("/fh/tempspace/gottardo_r/yezheng_working/meta/Histone_IgG_pairing.csv") %>% select(histone) %>% mutate(path = histone) %>% filter(! histone %in% blackList) %>% separate("histone", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") 
# fileListUni = fileList %>% select(Histone, CD, CellType, Experiment) %>% unique
# convertGR = function(peak){
#     peak.gr = GRanges(seqname = peak$V1, IRanges(peak$V2, peak$V3), strand = "*")
#     return(peak.gr)
# }
# 
# reprodD = c()
# for(i in 1:nrow(fileListUni)){
#   print(i)
#   hist = fileListUni[i, 1]
#   cd = fileListUni[i, 2]
#   cell = fileListUni[i, 3]
#   exp = fileListUni[i, 4]
#   hdList = fileList %>% filter(Histone == hist, CD == cd, CellType == cell, Experiment == exp)
#   for(p1 in 1:nrow(hdList)){
#       for(p2 in 1:nrow(hdList)){
#         if(p1 != p2){
#           
#           ## MACS2
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/stringent/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/stringent/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "MACS2") %>% rbind(reprodD)
#           
#            ## SEACR Control
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/stringent/SEACR/seacr_control.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/stringent/SEACR/seacr_control.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "SEACR_Control") %>% rbind(reprodD)
#     
#           p1N = length(peak1.gr)
#           p2N = length(peak2.gr)
#           peak1Con.gr = peak1.gr
#           peak2Con.gr = peak2.gr
#           
#           ## SEACR Top
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/stringent/SEACR/seacr_top0.01.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/stringent/SEACR/seacr_top0.01.peaks.stringent.bed")) %>% filter(V1 != "chrY", V1 != "chrM") %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "SEACR_Top") %>% rbind(reprodD)
#           
#           ## SEACR Control in Top
#           peak1ConTop.gr = peak1Con.gr[findOverlaps(peak1Con.gr, peak1.gr)@from %>% unique]
#           peak2ConTop.gr = peak2Con.gr[findOverlaps(peak2Con.gr, peak2.gr)@from %>% unique]
#           overlap = findOverlaps(peak1ConTop.gr, peak2ConTop.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1ConTop.gr) * 100, peakN = length(peak1ConTop.gr), caller = "SEACR_ControlInTop") %>% rbind(reprodD)
#           
#           ## MACS2 - topN
#           peak1.gr = fread(paste0(inPath, hdList$path[p1], "/peakCalling/stringent/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% top_n(p1N, wt = V9) %>% convertGR()
#           peak2.gr = fread(paste0(inPath, hdList$path[p2], "/peakCalling/stringent/MACS2/macs2_noControl_peaks.narrowPeak")) %>% filter(V1 != "chrY", V1 != "chrM") %>% top_n(p2N, wt = V9) %>% convertGR()
#           overlap = findOverlaps(peak1.gr, peak2.gr)
#           reprodD = data.frame(info = hdList$path[p1], overlapHD = hdList$HD[p2], overlapN = overlap@from %>% unique %>% length, overlapP = overlap@from %>% unique %>% length/length(peak1.gr) * 100, peakN = length(peak1.gr), caller = "MACS2(TopN=SEACR_Control)") %>% rbind(reprodD)
#          
#         }
#         
#       }
#     
#   }
#   
# }
# reprodD$caller = factor(reprodD$caller, levels = c("MACS2", "SEACR_Control", "SEACR_Top", "MACS2(TopN=SEACR_Control)", "SEACR_ControlInTop"))
# save(reprodD, file = paste0(outPath, "RData/reprodD_rmDuplicates.RData"))


load(file = paste0(outPath, "RData/reprodD_rmDuplicates.RData"))
```

#### 5.1 Variation of number of peaks among donors

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
reprodData = separate(reprodD, "info", into = c("Histone", "CD", "CellType", "Experiment", "HD", "Replicate"), sep = "_") %>% mutate(CellType = factor(toupper(CellType), levels = c("BULK", "N", "CM", "EM", "EMRA")))

reprodData %>% select(-Replicate, -overlapHD, -overlapN, -overlapP) %>% unique %>% group_by(Histone, CD, CellType, Experiment, caller) %>% summarize(sd = sd(peakN)) %>%
     filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = sd, fill = CellType)) +
    geom_boxplot() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~caller, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("Standard Deviation of Number of Peaks being Called") +
    xlab("") +
    ggtitle("Removing Duplicates")
```

Zoom in y-scale:0-10,000
```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
reprodData %>% select(-Replicate, -overlapHD, -overlapN, -overlapP) %>% unique %>% group_by(Histone, CD, CellType, Experiment, caller) %>% summarize(sd = sd(peakN)) %>%
     filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = sd, fill = CellType)) +
    geom_boxplot() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~caller, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("Standard Deviation of Number of Peaks being Called") +
    xlab("") +
    ggtitle("Removing Duplicates") +
    coord_cartesian(ylim = c(0, 10000))
```

#### 5.2 Peak overlapping among donors

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=11, fig.height=8}
reprodData %>% filter(CD == "CD8", CellType != "MAIT", CellType != "NA", Experiment != "PatientIP") %>% ggplot(aes(x = CellType, y = as.numeric(overlapP), fill = CellType)) +
    geom_boxplot() +
    # geom_jitter(aes(color = Experiment), position = position_jitter(0.15)) +
    facet_grid(Histone~caller, scales = "free", space = "free_x") +
    scale_fill_viridis(discrete = TRUE, begin = 0.1, option = "magma", alpha = 0.5) +
    scale_color_viridis(discrete = TRUE, begin = 0.1) +
    theme_bw(base_size = 12) +
    ylab("Reproducibility (%)") +
    xlab("") +
    ggtitle("Removing Duplicates")
```











